---
title: "Lab 2"
author: "Imani Russell"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

B. Attach packages 
```{r}
library(tidyverse)
library(janitor)
library(MASS)
library(ggrepel)
library(RColorBrewer)
library(factoextra)
library(ggbiplot)
library(shinythemes)
```

if you want to override (set a function manually throughout an entire workspace, you can assign the fxn name explicityly

##1. Ordinal logistic regression for political party affil by ethnicity, age, sex

```{r}
pol_party <- read_csv("pol_party.csv")
```

```{r}
names(pol_party)
```

```{r}
pol_party <- read_csv("pol_party.csv") %>%
  clean_names() %>%
  select(participant_id:read_carefully)

# A little more wrangling

pol_df <- pol_party %>%
  select(birth_year, gender, ethnicity, political party)
```

Note: MASS package also has a 'select' function that will mask 'select' in dplyr - we don't want that! There are two things we can do: 

(1) State the package explicitly that the function comes from when using it, e.g. dplyr::select

(2) Generally override that the default select is from MASS as follows: 

    select <- dplyr::select
    
```{r, message = FALSE}
select <- dplyr::select
```


###1. OLR with political party (Open Science Framework: https://osf.io/8n6s4/)

####a. Get data (pol_party.csv) and do some wrangling

```{r, message = FALSE}

pol_party <- read_csv("pol_party.csv") %>% 
  clean_names() # janitor function for snake case

pol_df <- pol_party %>% 
  select(birth_year, gender, ethnicity, political_party) %>% 
  filter(birth_year > 1900, political_party != "NA", political_party <= 7) %>% 
  mutate(age = 2016 - birth_year)

# Political party: 1 = very liberal, 7 = very conservative

# WHen doing OLR, we want our dependent variable to be of class ORDERED FACTOR

pol_df$political_party <- factor(pol_df$political_party, ordered = TRUE, levels = c("1", "2", "3", "4", "5", "6", "7"))

# update categorical predictor variables

pol_df$ethnicity <- factor(pol_df$ethnicity)
pol_df$gender <- factor(pol_df$gender)
```

Basic data exploration and visualization

```{r}
counts_eth <- pol_df %>%
  group_by(ethnicity, political_party) %>%
  tally()

counts_g <- pol_df %>%
  group_by(gender, political_party) %>%
  tally()

ggplot(pol_df, aes(x = political_party)) +
  geom_histogram(aes(fill = ethnicity), stat = "count") +
  scale_fill_brewer(palette = "YlGnBu") +
  facet_wrap(~ethnicity, scales = "free")

```

# THESE DATA ARE PRECARIOUS AND N IS TOO SMALL!!! FOR GROUPS

```{r}

# THIS IS NOT PUBLISHABLE DATA

pol_model <- polr(political_party ~ age + gender + ethnicity, data = pol_df)

summary(pol_model)

# coeffs are in log-odds
# if log odds coeff of 0 = no effect --> age value is VERY SMALL - as age increases, expect to see a slight increase of people ranking themselves higher on that scale, but it's super tiny

# for factors, always have a reference level
# males have increase of log odds relative to females
# males more likely to rank higher on liberal to conservative scale

# ethnicity all compared to reference of asian
# native american has no meaning because only 3

# easier to just look at ODDS

exp(coef(pol_model))

# if odds = 1 means no change in the odds
# age has no effect on party affiliation
# increase in odds of males selecting higher ranks compared to females of similar age and ethnicity

### 

# Once we've decided on a model, we can always make predictions using it

predictions <- predict(pol_model, type = "probs")

df <- data.frame(pol_df, predictions)

# gives breakdown on model predictions for each individual
```

###2. Principal components analysis (PCA)

Using WorldBank environmental indicators

```{r}

wb_10 <- read_csv("wb_10.csv") %>% 
  select(-X1) %>% 
  column_to_rownames('code')
  
wb_pca <- prcomp(wb_10[3:8], scale = TRUE)
  
summary(wb_pca)

# ggbiplot for visualization
ggbiplot(wb_pca, groups = wb_10$region) +
  geom_text_repel(label = rownames(wb_10), size = 2, segment.size = 0.2) +
  theme_bw() +
  scale_x_continuous(limits = c(-6,3)) +
  scale_y_continuous(limits = c(-6,4))

# factoextra package (built to help visualize outcomes of ordination methods)

# screeplot:

fviz_screeplot(wb_pca)

# see contributions of variables to different components:

fviz_contrib(wb_pca, choice = "var", axes = 1, top = 6)

# Biplot showing points and vectors:

fviz_pca_biplot(wb_pca,
                col.var = "cos2",
                label = "var",
                repel = TRUE)
```


































